### 2010-02-16 doc/tech_doc/a-rex/arex_tech_doc.tex
add explanation of download checks and retry strategy, and improve formatting

\begin{itemize}
\item Download is initiated by the A-REX -- This is the case for files defined
in the JD (with name and source). The A-REX alone is responsible to
ensure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-REX fully relies on HED framework~\cite{hed} for data transferring
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-REX. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-REX).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
of these files. They have to be specified in the JD. If the job fails
during any stage of processing, no attempt is made to transfer those
files to their final destination, unless the option \emph{preserve=yes}
is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}
All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from first group are transferred to
second one.

@@

\begin{itemize}
\item \textbf{Accepted} -- In this state the job has been submitted to a
CE but is not processed yet. The A-REX will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} -- The input data is being gathered in the SD (stage-in).
The A-REX is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. 
If this limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submitting} -- The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The corresponding
backends for many LRMSs are provided with the default installation.
If the local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the aggregate number of jobs in \textbf{Submitting} and \textbf{Executing}
states.
\item \textbf{Executing} -- The job is queued or being executed in the LRMS.
The A-REX takes no actions except waiting until the job finishes.
\item \textbf{Killing} -- Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} -- The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state.It is possible to limit the number of simultaneous jobs in this state.
\item \textbf{Finished} -- No more processing is performed by the A-REX.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Information Interface of the A-REX. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} -- The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

>>>>

\item Download is initiated by the A-REX -- This is the case for files
  defined in the JD (with name and source). The A-REX alone is
  responsible to ensure that all required files will be available in
  the SD.\\ The supported protocols for sources at the moment are (in
  case of full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over
  SSLv3). Some less standard sources and \emph{Indexing Services} are
  also supported, these are described in Section~\ref{sec:urls}. The
  A-REX fully relies on the HED framework~\cite{hed} for data
  transferring capabilities and so the actual set of supported
  protocols depends on the installed Data Management Components of the
  HED.

\item Upload is initiated by the user directly or through the User
  Interface (UI). Because the SD becomes available immediately at the
  time of submission of the JD, the UI can (and should) use that to
  upload data files which are not otherwise accessible by the
  A-REX. Examples of such files are the main executable of the job,
  the job's input files, etc.  These files can (and should) also be
  specified in the JD.
\end{itemize}

\textbf{There is no} other reliable way for a job to obtain input data
on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and any
other remote data transport during execution of a job is not
guaranteed (at least not by A-REX).

At the start of a download, a dummy file with the same name and size
as the source file is created in the SD to check that enough space is
available. As the transfer proceeds this dummy file is over-written
with the real data. If the file is to be cached (see
Section~\ref{sec:cache}) then the pre-allocation of space is not done,
as cache space is managed by the A-REX. In the case of caching two
extra validation checks are also performed: firstly, if the source is
an Indexing Service then metadata (file size, checksum) reported by
the Indexing Service and the service hosting the physical replica are
compared, if available. If they differ then that replica is not
downloaded. Secondly, on completion of the download the file size is
compared to that reported by the source - if they differ then the
download is failed.

@@

\item Files which are supposed to be moved to a \emph{Storage Element}
  (SE) and optionally registered in some Indexing Service like the
  Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
  of these files. They have to be specified in the JD. If the job
  fails during any stage of processing, no attempt is made to transfer
  those files to their final destination, unless the option
  \emph{preserve=yes} is specified in their URLs.

\item Files which are supposed to be fetched by the user -- The user
  has to use a tool like the UI to obtain these files. They
  \textbf{must} also be specified in the JD.
\end{itemize}

All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from first group are transferred to
second one.

@@

\item \textbf{Accepted} -- In this state the job has been submitted to
  a CE but is not processed yet. The A-REX will analyze the JD and
  move to the next stage. If the JD can not be processed the job will
  be canceled and moved to the state \textbf{Finishing}.

\item \textbf{Preparing} -- The input data is being gathered in the SD
  (stage-in).  The A-REX is downloading the files specified in the JD
  and is waiting for files which are supposed to be uploaded by the
  UI. If all files are successfully gathered the job moves to the next
  state. If \textbf{any} file can't be downloaded or it takes the UI
  too long to upload a file, the job moves to \textbf{Finishing}
  state. It is possible to put a limit on the number of simultaneous
  \textbf{Preparing} jobs.  If this limit is exceeded, jobs ready to
  enter the \textbf{Preparing} state will stay in the
  \textbf{Accepted} state, but prefixed with the PENDING:
  mark. Exceptions are jobs which have no files to be
  downloaded. These are processed out of limits. If the A-REX fails to
  download a file due to a temporary error with a remote service, the
  job will be moved back to the \textbf{Accepted} state, but prefixed
  with the PENDING: mark, and will wait in that state for some time
  before being retried. The waiting time increases exponentially with
  each attempt.

\item \textbf{Submitting} -- The job is being passed for execution to
  the \emph{Local Resource Management System} (LRMS). The
  corresponding backends for many LRMSs are provided with the default
  installation.  If the local job submission is successful the job
  moves to the \textbf{Executing} state. Otherwise it moves to
  \textbf{Finishing}. It is possible to limit the aggregate number of
  jobs in \textbf{Submitting} and \textbf{Executing} states.

\item \textbf{Executing} -- The job is queued or being executed in the
  LRMS.  The A-REX takes no actions except waiting until the job
  finishes.

\item \textbf{Killing} -- Necessary action to cancel the job in the
  LRMS is being taken.

\item \textbf{Finishing} -- The output data is being processed
  (stage-out).  Specified data files are moved to the specified SEs
  and are optionally registered at an Indexing Service. The user can
  download data files from the SD by using the UI or other adequate
  tool. All the files not specified as output files are removed from
  the SD at very beginning of this state. It is possible to limit the
  number of simultaneous jobs in this state.  If the A-REX fails to
  upload a file due to a temporary error with a remote service, the
  job will be moved back to the \textbf{Executing} state, but prefixed
  with the PENDING: mark, and will wait in that state for some time
  before being retried. The waiting time increases exponentially with
  each attempt.

\item \textbf{Finished} -- No more processing is performed by the
  A-REX.  The user can continue to download data files from the
  SD. The SD is kept available for some time (default is 1
  week). After that the job is moved to the state
  \textbf{Deleted}. The 'deletion' time can be obtained by querying
  the Information Interface of the A-REX. If a job was moved to
  \textbf{Finished} because of failure, it may be restarted on request
  of a client. When restarted, a job is moved to the state previous to
  the one in which it failed and is assigned mark PENDING.  This is
  needed in order to not break the configuration limits. Exception is
  a job failed in \textbf{Executing} state and lacking input files
  specified in JD. Such a job is treated like failed in
  \textbf{Preparing} state.

\item \textbf{Deleted} -- The job is moved to this state if the user
  have not requested job to be cleaned before the SD's lifetime
  expires.  Only minimal subset of information about such job is
  kept. The SD is not available anymore.

In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by the
  user}. No files will be moved to their destination SE.
